Toward Supervised Reinforcement Learning with Partial States for Social HRI

### 2017-08-22 paper.tex
Early updates based on reviews

Approaches from animal conditioning have also been transfered to Machine
Learning. For example, the TAMER framework~\cite{knox2009interactively} is an
adaptation of ``shaping'', tuning an animal's behaviour by providing rewards to
guide it to the desired behaviour. Unlike other RL setups, with TAMER, the
reward is not provided by the environment but by a human observer. With this
approach the teacher can adapt the rewards to the current policy and performance
of the agent, for example starting to positively reward suboptimal behaviours
that are better than the current policy and later, as the agent improves its
behaviour, reward these same suboptimal behaviours negatively and only the
optimal ones positively.

@@

SPARC gives control of the robot's actions to an expert, who can guide the
exploration in the desired part of the environment which ensures the robot's
behaviour is consistently appropriate. As the exploration is guided, and all the
actions are useful, the learning can be faster than autonomous learning or
feedback based learning as illustrated in Figure~\ref{fig:comparison}.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{./figs/motivation.pdf}
    \caption{An illustration of the evolution over time of the performance,
    autonomy and human workload for an autonomous learner, an approach using
    human rewards, and SPARC.}
    \label{fig:comparison}
\end{figure}

In~\cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot, making the
assumption that every action has been explicitly or implicitely approved by the
supervisor. However this method was directly mapping a single (state, action)
pair to a reward without making use of any kind of generalisation. As such it
was not applicable to environments with a continuous or high dimensional space
or non-deterministic transitions from one state to another, elements which are
typically present in social interactions.

@@

To make RL applicable in high dimensional or continuous states, the algorithm
requires a way to generalise knowledge to unseen states. A classic approach is
to use a feed-forward neural network or to use deep reinforcement learning
relying on deep neural networks to learn a the value function. These approaches
rely on having a large number of datapoints to converge toward a good function
approximator. However, in many applications and especially in HRI, these amounts
of data are not available or obtainable due to practical constraints.
Furthermore, human responses are often noisy and lack consistency, so methods
are needed which can generalise from a low number of noisy data points.

@@

As the supervisor can estimate the future impacts of an action, the problem of
credit assignment for delayed rewards can be ignored which allows us to consider
only a myopic approach.

@@

\begin{algorithm}
    \DontPrintSemicolon
    \SetKwInOut{Input}{inputs}\SetKwInOut{Output}{output}
    \Input{Current state $s$, set of  $(a,s',r)$}
    \Output{selected action $\pi(s)$}
    \ForEach{$a \in \mathcal{A}$}{
        \ForEach{$p=(s',r) \in \mathcal{C}_{a}$}{
            compute similarity $\Delta$ between $s$ and $s'$:
            $\Delta(p)=1-\frac{\sum_{i}^{n'}(s'(i)-s(i))^{2}}{n'}$
        }
        compute expected reward $\hat{r}(a)$ for taking $a$ in $s$:
        $\hat{r}(a) = max(\Delta) \cdot r(arg\, max_{p} \Delta(p))$\\
        with $r(\Delta(p))$ the reward of $p = (s',r)$ 
    }
    Select the action with the maximum expected reward:
    $\pi(s) = arg\, max_{a} \hat{r}(a)$

@@

When facing a new state $s$ where an action has to be selected, the agent can
select an action following Algorithm~\ref{algo}. For each action $a \in
\mathcal{A}$, we take the pair $(s',r)$ with the closest $s'$ to the current
state (as defined by normalised distance over the dimensions where $s'$ is
defined). That way, each action $a$ can be associated to an expected reward
defined by the product between the similarity of the closest partial state known
for $a$ and the reward obtained for executing $a$ in that partial state.
Finally, the action with the highest expected reward can be selected.

>>>>

Approaches from animal conditioning have also been transfered to Machine
Learning. For example, the TAMER framework~\cite{knox2009interactively} is an
adaptation of ``shaping'', tuning an animal's behaviour by providing rewards to
guide it to the desired behaviour. Unlike other RL setups, with TAMER, the
reward is not provided by the environment but by a human observer. TAMER derives
a reward function from the human feedbacks and uses this function to evaluate
the current behaviour. In a similar approach, MacGlashan et al. proposed COACH
in~\cite{macglashan2017interactive}. COACH assumes that human rewards represent
the advantage function, i.e. how much the current action is better than the
current policy. With COACH, the teacher can adapt the rewards to the current
policy and performance of the agent, for example starting to positively reward
suboptimal behaviours that are better than the current policy and later, as the
agent improves its behaviour, reward these same suboptimal behaviours negatively
and only the optimal ones positively.

@@

SPARC gives control of the robot's actions to an expert, who can guide the
exploration in the desired part of the environment which ensures the robot's
behaviour is consistently appropriate. As the exploration is guided, and all the
actions are useful, the learning can be faster than autonomous learning or
learning based on human feedback as illustrated in Figure~\ref{fig:comparison}.
In Autonomous Learning, the agent has first to discover its environment to start
gathering relevant feedbacks, leading to a low initial performance and a slow
improvement in early stage of learning. With a teaching based on human
feedbacks, the teacher can quickly provide feedbacks on actions to reach more
quickly an efficient policy. However, as the teacher only provide feedbacks and
cannot prevent the agent to make mistakes, the initial performance can be poor.
On the other hand, SPARC, with the control of the teacher over the executed
actions can prevent the agent to make mistakes in early stages, achieving a high
performance even at the start of the learning. The total blending between
autonomous execution of actions and demonstrations and the control over
\emph{every} actions executed by the agent are the specificities of SPARC
compared to other approaches using humans to guide the learning of an agent.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{./figs/motivation.pdf}
    \caption{An illustration of the evolution over time of the performance,
    autonomy and human workload for an autonomous learner, an approach using
    human feedbacks, and SPARC.}
    \label{fig:comparison}
\end{figure}

In~\cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot, making the
assumption that every action has been explicitly or implicitely approved by the
supervisor. However this method was directly mapping a single (state, action)
pair to a reward without making use of any kind of generalisation. As such it
was not applicable to environments with a continuous or high dimensional space
or non-deterministic transitions from one state to another, elements which are
typically present in social interactions. Similarly to TAMER or COACH, this
approach allows to reproduce an action policy from a teacher even in the absence
of a reward function, but as the teacher has control over the robot actions, the
learning can be faster and safer.

@@

To make RL applicable in high dimensional or continuous states, the algorithm
requires a way to generalise knowledge to unseen states. A classic approach is
to use a feed-forward neural network or deep learning to learn a the value
function generalising to unseen states. Neural network rely on having a large
number of datapoints to converge toward a good function approximator.
Alternatively, Cobo et al. proposed in~\cite{cobo2011automatic} to abstract
features automaticly from demonstrations to learn faster. But even in this case,
the number of datapoints required is still high (around 1000 samples per
participants). However, in many applications and especially in HRI, these
amounts of data are not available or obtainable due to practical constraints.
Furthermore, human responses are often noisy and lack consistency, so methods
are needed which can generalise from a low number of noisy data points.

@@

As the supervisor can estimate the future impacts of an action, the problem of
credit assignment for delayed rewards can be ignored which allows us to consider
only a myopic approach in a fashion similar to TAMER~\cite{knox2009interactively}.

@@

\begin{algorithm}
    \DontPrintSemicolon
    \SetKwInOut{Input}{inputs}\SetKwInOut{Output}{output}
    \Input{Current state $s$, set of  $(a,s',r)$}
    \Output{selected action $\pi(s)$}
    \ForEach{$a \in \mathcal{A}$}{
        \ForEach{$p=(s',r) \in \mathcal{C}_{a}$}{
            compute similarity $\Delta$ between $s$ and $s'$:
            $\Delta(p)=1-\frac{\sum_{i}^{n'}(s'(i)-s(i))^{2}}{n'}$
        }
        find closest pair $\hat{p}$:\\
        $\hat{p} = arg\, max_{p} \Delta(p)$\\
        compute expected reward $\hat{r}(a)$ for taking $a$ in $s$:\\
        $\hat{r}(a) = \Delta (\hat{p}) \cdot r(\hat{p})$\\
        with $r(p)$ the reward of $p = (s',r)$ 
    }
    Select the action with the maximum expected reward:
    $\pi(s) = arg\, max_{a} \hat{r}(a)$

@@

When facing a new state $s$ where an action has to be selected, the agent can
select an action following Algorithm~\ref{algo} in a instance-based learning
fashion. For each action $a \in \mathcal{A}$, we take the pair $(s',r)$ with the
closest $s'$ to the current state (as defined by the average quadratic distance
over the normalised dimensions where $s'$ is defined). That way, each action $a$
can be associated to an expected reward defined by the product between the
similarity of the closest partial state known for $a$ and the reward obtained
for executing $a$ in that partial state.  Finally, the action with the highest
expected reward can be selected.

The normalisation of each dimension of the state allows distances to be
comparable as values on all dimensions have the same range. Additionally taking
the average quadratic distance over each defined dimension allows to compare
distances even when states are defined on a different number of dimensions.
